---
title: "Lendingclub"
author: "elsa"
date: "8 septembre 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
```

# Setup
```{r}
# unnamed chunk (all chunks are not named in fact)
packages <- c("tidyverse", "stringr", "tidytext", "fiftystater", 
              "data.table", "DescTools", "wordcloud", "lubridate","RColorBrewer", "maps", "corrplot", "mapproj") # line too long => 80 characters max !


for (package in packages) {
    if (!require(package, character.only=T, quietly=T)) {
        install.packages(package, dependencies = TRUE)
        library(package, character.only=T)
    } else {
        library(package, character.only=T)
    }
}

```



```{r}
data_lending <- read_csv("../inputs/loan.csv", 
  col_types = cols(
  id = col_character(),
  member_id = col_character(),
  loan_amnt = col_double(),
  funded_amnt = col_double(),
  funded_amnt_inv = col_double(),
  term = col_character(),
  int_rate = col_double(),
  installment = col_double(),
  grade = col_character(),
  sub_grade = col_character(),
  emp_title = col_character(),
  emp_length = col_character(),
  home_ownership = col_character(),
  annual_inc = col_double(),
  verification_status = col_character(),
  issue_d = col_character(),
  loan_status = col_character(),
  pymnt_plan = col_character(),
  url = col_character(),
  desc = col_character(),
  purpose = col_character(),
  title = col_character(),
  zip_code = col_character(),
  addr_state = col_character(),
  dti = col_double(),
  delinq_2yrs = col_double(),
  earliest_cr_line = col_character(),
  inq_last_6mths = col_double(),
  mths_since_last_delinq = col_double(),
  mths_since_last_record = col_double(),
  open_acc = col_double(),
  pub_rec = col_double(),
  revol_bal = col_double(),
  revol_util = col_double(),
  total_acc = col_double(),
  initial_list_status = col_character(),
  out_prncp = col_double(),
  out_prncp_inv = col_double(),
  total_pymnt = col_double(),
  total_pymnt_inv = col_double(),
  total_rec_prncp = col_double(),
  total_rec_int = col_double(),
  total_rec_late_fee = col_double(),
  recoveries = col_double(),
  collection_recovery_fee = col_double(),
  last_pymnt_d = col_character(),
  last_pymnt_amnt = col_double(),
  next_pymnt_d = col_character(),
  last_credit_pull_d = col_character(),
  collections_12_mths_ex_med = col_double(),
  mths_since_last_major_derog = col_character(),
  policy_code = col_double(),
  application_type = col_character(),
  annual_inc_joint = col_character(),
  dti_joint = col_character(),
  verification_status_joint = col_character(),
  acc_now_delinq = col_double(),
  tot_coll_amt = col_character(),
  tot_cur_bal = col_character(),
  open_acc_6m = col_character(),
  open_il_6m = col_character(),
  open_il_12m = col_character(),
  open_il_24m = col_character(),
  mths_since_rcnt_il = col_character(),
  total_bal_il = col_character(),
  il_util = col_character(),
  open_rv_12m = col_character(),
  open_rv_24m = col_character(),
  max_bal_bc = col_character(),
  all_util = col_character(),
  total_rev_hi_lim = col_double(),
  inq_fi = col_character(),
  total_cu_tl = col_character(),
  inq_last_12m = col_character()
))
```

# Exploratory Data Analysis

```{r}
# is it relevant ?
summary(data_lending)
```

## Missing values

```{r}
# DT::datable makes nice tables to PRINT, why load it in a variable ?
missing_table <-  data_lending %>% 
  map_dbl(~ round(sum(is.na(.x))/length(.x), digits=2)) %>%
  data_frame(a=names(.), b=.) %>% # a and b as column names ?
  DT::datatable()

delete_missing <- missing_table$x$data %>%  filter(b>0.7) %>% select(a) %>% unlist() #too long, and too complex because you manipulate a DT::datatable object
# why 0.7 ?

data_lending_deleted_missing <- data_lending[ , -which(names(data_lending) %in% delete_missing)] # can you use select ? select_if here ? too long

missing_table
# in the end you removed all columns with too many NAs but didnt write it
```

## Feature Cleaning

```{r}
#extract numbers from term
data_clean <- data_lending_deleted_missing %>% 
  mutate(term= as.numeric(str_extract_all
                          (term, "[[:digit:]]+"))) # format !
# why str_extract_all vs str_extract ?

#extract numbers from employment length
data_clean <- data_clean %>% 
  mutate(emp_length= as.numeric(str_extract_all
                                (emp_length, "[[:digit:]]+"))) # format !
# why str_extract_all vs str_extract ?

#Binary Verification Status
data_clean <- data_clean %>% 
  mutate(verification_status= ifelse # format !
         (verification_status=="Not Verified", 0, 1))
# here we binarize but 3 levels !

#Issue Date into date format

months_dict <- c("01", "02", "03", "4", "5", "6", "7", "8", "9", "10", "11", "12") # too long
names(months_dict) <- c("Jan", "Feb", "Mar", "Apr", "May",
                        "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

data_clean <- data_clean %>% 
  separate(issue_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(issue_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(issue_d= dmy(issue_d)) %>% 
  select(-m_d)
# better naming ?

#Earliest Credit Line into date format
data_clean <- data_clean %>% 
  separate(earliest_cr_line, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(earliest_cr_line, d_d, good_m, y_d, sep="-") %>% 
  mutate(earliest_cr_line= dmy(earliest_cr_line)) %>% 
  select(-m_d)
# better naming ?

# Last payment into date format
  
data_clean <- data_clean %>% 
  separate(last_pymnt_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_pymnt_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_pymnt_d= dmy(last_pymnt_d)) %>% 
  select(-m_d)
# better naming ? why is this one much longer to execute ?
# why failed to parse messages ?

# Last Credit Pull into date format
data_clean <- data_clean %>% 
  separate(last_credit_pull_d, c("m_d", "y_d"), sep = "-") %>% 
  mutate(d_d = "01") %>% 
  mutate(good_m = months_dict[m_d]) %>% 
  unite(last_credit_pull_d, d_d, good_m, y_d, sep="-") %>% 
  mutate(last_credit_pull_d= dmy(last_credit_pull_d)) %>% 
  select(-m_d)
# better naming ?

# filtering non-individual loans 
# Why ?
data_clean <- data_clean %>% 
  filter(application_type=="INDIVIDUAL") %>% 
  select(-application_type)
  
```

# General charactéristics of the business and the loans
MAG : ortho + can we specify row count inline ?
The data set contains now 886868 observations and 53 variables.

## Evolution of the Lending Club business

```{r}
amnt_df <- data_clean %>% 
  select(issue_d, loan_amnt) %>% 
  group_by(issue_d) %>% 
  summarise(Amount = sum(loan_amnt))
# why an intermediary dataset amnt_df ?

 ggplot(amnt_df,aes(x = issue_d, y = Amount)) +
  geom_line() + 
  labs(x="Date issued", y= "Loan Amount", title="Evolution of amount required in time")
# format ? theme ?
# does the fake 01/ in issue_d implies something on the graph ?
```

## Geographical analysis in Loan Volume

```{r}
data_state <- data_clean %>% 
  mutate(Abbreviation = addr_state)
# renaming ? if so do not use mutate use rename

state_by_volume <- data_state %>% 
  group_by(Abbreviation) %>%
  summarise(value = n())
# use count() idiom

full_name_state <- read_csv("./resources/full_name_state.csv")

state_by_volume <-  state_by_volume %>% inner_join(full_name_state) %>% # key for the join ?
  select(-Abbreviation) %>% 
  mutate(region=State) %>% 
  select(-State)

full_region_vec <- rownames(state.x77)

full_region_df <- as.data.frame(full_region_vec) # vs as_data_frame ?
colnames(full_region_df) <- "region" # necessary ?

state_by_volume <-  state_by_volume %>% 
  inner_join(full_region_df) %>% # key ?
  mutate(value= as.numeric(value))

state_by_volume <- state_by_volume %>%
  mutate(region = str_to_lower(region))

map_volume <- ggplot(state_by_volume, aes(map_id = region)) + 
  # map points to the fifty_states shape data
  geom_map(aes(fill = value), map = fiftystater::fifty_states) + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "", title="Loan number by State")
 theme(legend.position = "bottom", 
     panel.background = element_blank())

map_volume

```

## Geographical analysis in Loan Value

```{r}
data_state <- data_clean %>% 
  mutate(Abbreviation = addr_state)

state_by_value <- data_state %>% 
  group_by(Abbreviation) %>%
  summarise(value = sum(loan_amnt, na.rm=TRUE))

full_name_state <- read_csv("./resources/full_name_state.csv")

state_by_value <-  state_by_value %>% inner_join(full_name_state) %>% 
  select(-Abbreviation) %>% 
  mutate(region=State) %>% 
  select(-State)

full_region_df <- as.data.frame(full_region_vec)
colnames(full_region_df) <- "region"

state_by_value <-  state_by_value %>% 
  inner_join(full_region_df) %>% 
  mutate(value= as.numeric(value))
state_by_value <- state_by_value %>% mutate(region = str_to_lower(region))
map_value <- ggplot(state_by_value, aes(map_id = region)) + 
  # map points to the fifty_states shape data
  geom_map(aes(fill = value), map = fiftystater::fifty_states) + 
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  labs(x = "", y = "", title="Loan Amount by State")
 theme(legend.position = "bottom", 
     panel.background = element_blank())

map_value

```


## Distribution of loan amount 

```{r}
Desc(data_clean$loan_amnt, main = "Loan amount distribution", plotit = FALSE) 
# relevant ?
```


## Distribution of loan amounts according to the length 

```{r}
data_term <- data_clean %>% 
  mutate(term=as.factor(term)) # vs term = factor(term) ?

box_amnt <- ggplot(data_term, aes(term, loan_amnt))
# why intermediary box_amnt object ?
box_amnt + geom_boxplot(aes(fill = term)) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by term",
    x = "Term",
    y = "Amount")
# theme ?
```


## Distribution of Incomes 

```{r}
ggplot(data_clean, aes(x=annual_inc)) +
  geom_histogram(bins = 100) + # why 100 ?
  xlim(0, 400000) + #why set xlim
  labs(x= "Annual Income",
       title= "Distribution of Annual Income")
# theme ? labs ?
```


## Terms compared to incomes

```{r}
data_term <- data_clean %>% 
  mutate(term=as.factor(term)) # vs term = factor(term) ?
# the data_term object has already been created

box_income <- ggplot(data_term, aes(term, annual_inc))
box_income + geom_boxplot(aes(fill = term)) +
  theme(axis.text.x = element_blank()) +
  ylim(0, 400000 ) + # why ?
  labs(title = "Income by term",
    x = "Term",
    y = "Income")
# theme ?
```

## Loan amount compared to income

```{r}
data_proportion <- data_term %>% 
  filter(annual_inc<500000) %>% # format
  filter(loan_status!="Current") %>% 
  mutate(prop_inc_amnt= loan_amnt/annual_inc)
# naming ?

ggplot(data_proportion, aes(x=prop_inc_amnt)) +
  geom_histogram(bins = 50)+
  facet_grid(.~ term)+
  xlim(0, 0.7)
# pipe data in ggplot

data_36 <- data_proportion %>% 
  filter(term==36)
# naming
data_60 <- data_proportion %>% 
  filter(term==60)
# naming

# 36 terms
ggplot(data_36, aes(x=annual_inc, y= loan_amnt)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 0.50, color="red") +
  labs(x= "Annual Income", y="Loan Amount", title="Loan amount for 36 terms compared to Annual Income")
# theme ? format ?
# where does the abline come from ?

#60 terms
ggplot(data_60, aes(x=annual_inc, y= loan_amnt)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 0.52, color="red")+
  labs(x= "Annual Income", y="Loan Amount", title="Loan amount for 60 terms compared to Annual Income ")

```



## Est-ce que les gens qui empruntent beaucoup par rapport à leur salaire sont plus en défaut?

```{r}
data_proportion_factor <- data_proportion %>% 
  mutate(term=as.factor(term)) %>% # vs factor
  mutate(loan_status= as.factor(loan_status)) # vs factor

ggplot(data_proportion_factor, aes(x=loan_status, y=prop_inc_amnt, fill=loan_status)) +
  geom_boxplot() +
  ylim(0,1) +
  facet_grid(.~term)
# theme ?
```

## ANOVA

```{r}
data_model_36 <- data_proportion_factor %>% 
  filter(term==36) %>% 
  filter(loan_status %in% c("Fully Paid", "Charged Off"))

model_proportion <- lm(prop_inc_amnt ~ loan_status, data=data_model_36)
summary(model_proportion)
anova(model_proportion)
confint(model_proportion)
# what's the idea ?

data_model_60 <- data_proportion_factor %>% 
  filter(term==60)%>% 
  filter(loan_status %in% c("Fully Paid", "Charged Off"))


model_proportion <- lm(prop_inc_amnt ~ loan_status, data=data_model_60)
summary(model_proportion)
anova(model_proportion)
confint(model_proportion)

```


# Purpose analysis

```{r}
Desc(data_clean$purpose, plotit = TRUE)
# relevant ?
```

```{r}
data_purpose <- data_term %>% 
  mutate(purpose= as.factor(purpose)) # vs factor ?

box_purpose <- ggplot(data_purpose, aes(purpose, loan_amnt))
# why intermediary box_purpose obj ?
box_purpose + geom_boxplot(aes(fill = purpose)) +
  theme(axis.text.x = element_blank()) +
  labs(title = "Loan amount by purpose",
    x = "term",
    y = "Amount")

```

```{r}
word_purpose <- data_clean %>% count(purpose, sort=TRUE) %>% 
  mutate(freq= n/nrow(data_clean))

wordcloud(word_purpose$purpose, word_purpose$freq, random.order = FALSE,
          rot.per = 0.3, min.freq = 1, use.r.layout=FALSE, colors=brewer.pal(8, "Dark2"))
# relevant ?

```



# ROI analysis

## Filtering current and non-finished loans

```{r}
data_roi_analysis <- data_clean %>%
  filter(loan_status %in% c("Charged Off", "Default", "Fully Paid", "In Grace Period", "Late (16-30 days)", "Late (31-120 days)")) %>% # format
  filter(term==36 & issue_d<ymd("2012-06-01") | term==60 & issue_d<ymd("2010-06-01")) %>% 
  filter(loan_amnt==funded_amnt)
  

```


## Verification of mathematical formulas and Calculation of actual benefits for the investors

```{r}
# Formulas to calclulate monthly payment due 
data_roi_analysis %>% 
  mutate(monthly_pymt_calculated= 
           (loan_amnt*(int_rate*0.01/12))/(1 - (1 +int_rate*0.01/12) ^(-term))) %>%         mutate(diff_pmt=monthly_pymt_calculated - installment) %>% 
  select(installment, monthly_pymt_calculated, diff_pmt)
# format ? can we do clearer ?

# Theoretical and Real total benefits

data_benef <- data_roi_analysis %>%
  mutate(theo_pymt = installment*term) %>% 
  mutate(theo_benef = theo_pymt - loan_amnt) %>% 
  mutate(real_benef= total_pymnt - loan_amnt)

# Investor theoretical and actual ROI

data_roi <- data_benef %>% 
  mutate(theo_roi = (theo_benef/loan_amnt)) %>% 
  mutate(diff_pymt = total_pymnt - total_pymnt_inv) %>% 
  mutate(real_roi= real_benef/loan_amnt) %>% 
  mutate(real_annual_roi= (real_roi + 1)^ (12/term) -1) %>% 
  mutate(theo_annual_roi= (theo_roi + 1)^ (12/term) -1)



```



## Theoretical Roi compared to actual status of loan 

```{r}
data_roi_status <- data_roi %>% 
  mutate(loan_status= as.factor(loan_status))

ggplot(data_roi_status, aes(x= loan_status, y=theo_roi)) +
  geom_boxplot(aes(fill=loan_status)) +
  theme(legend.position= "right") +
  labs(x="Loan Status", y="Theoretical ROI", title= "Theoretical ROI on loan compared to Loan Status")

```



## Theoretical ROI compared to Grades

```{r}
data_roi_status <- data_roi %>% 
  mutate(grade= as.factor(grade))

ggplot(data_roi_status, aes(x= grade, y=theo_roi)) +
  geom_boxplot(aes(fill=grade)) +
  theme(legend.position= "right") +
  labs(x="Grades", y="Theoretical ROI", title= "Theoretical ROI on loan compared to Grades")

```


## Theoretical ROI vs Actual ROI compared to Status

```{r}


ggplot(data_roi_status, aes(x= theo_roi, y = real_roi, color=loan_status)) +
  geom_point() +
  labs(x= "Theoretical ROI", y= "Actual ROI", title="Actual vs Theoretical Return on Investment") +
  geom_smooth(method = "lm", formula = y ~x, SE=FALSE)

```

## Theoretical ROI vs Actual ROI compared to Grades


```{r}
ggplot(data_roi_status, aes(x= theo_roi, y = real_roi, color=grade)) +
  geom_point() +
  labs(x= "Theoretical ROI", y= "Actual ROI", title="Actual vs Theoretical Return on Investment")


```

## Theoretical ROI vs Actual ROI compared to Emission Date of the loan

```{r}
ggplot(data_roi_status, aes(x= theo_roi, y = real_roi, color=issue_d)) +
  geom_point() +
  labs(x= "Theoretical ROI", y= "Actual ROI", title="Actual vs Theoretical Return on Investment")
```


## Mean Theoretical ROI vs Real ROI  

```{r}

mean(data_roi$theo_roi)
mean(data_roi$real_roi)

```

## Roi difference

```{r}
data_roi <- data_roi %>% 
  mutate(diff_roi = real_roi - theo_roi)

```

## Distribution of the difference

```{r}
ggplot(data_roi, aes(x=diff_roi, fill=loan_status)) +
  geom_histogram(bins=200)

```


```{r}

ggplot(data_roi, aes(x=loan_status, y=diff_roi, fill=loan_status)) +
  geom_boxplot() +
  facet_grid(.~grade)
```

## What default loan have positive difference between theoretical and actual roi and Why? 
We can notice quite high recovery fees

```{r}
  data_roi %>% filter(diff_roi>0.01, loan_status=="Charged Off") %>% DT::datatable()

```

## Late payments 

```{r}
data_roi_late <- data_roi %>% 
  mutate(loan_end= issue_d + term*30.4375) %>% # ??
  mutate(late_days= loan_end - last_pymnt_d)

fully_paid <- data_roi_late %>% 
  filter(loan_status=="Fully Paid")

ggplot(fully_paid, aes(x=round(as.numeric(late_days)/30.4375))) +
  geom_histogram(bins = max(fully_paid$late_days)) # ??


```

Correlation matrix

```{r}
data_roi_cor <- data_roi %>% 
  select_if(is_double) %>% 
  select_if(function(col) !is.Date(col)) %>% 
  mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .))) %>% 
  select(-collections_12_mths_ex_med, -policy_code, acc_now_delinq,
         -total_rev_hi_lim, -acc_now_delinq, -out_prncp, -out_prncp_inv) %>% 
  mutate()


cor_quanti <- cor(data_roi_cor, use="everything")
corrplot(cor_quanti, order="AOE", type="lower")

"Subset Matrix with correlations higher than 0.5"
df_quanti <- as.data.frame(cor_quanti)
subset_matrix <- df_quanti %>%
  mutate(var_name = rownames(df_quanti)) %>% 
  gather(var_name_2, corr_value, - var_name) %>% 
  filter(abs(corr_value) > 0.5, var_name != var_name_2) %>%
  select(var_name) %>% 
  distinct()
cor_subset <- cor_quanti[subset_matrix$var_name, subset_matrix$var_name]
corrplot(cor_subset, order="AOE", type="lower")
# package something ?

```



```{r}
ggplot(data_roi, aes(x=real_roi))+
  geom_histogram(bins=110)

```
# Probability

## Coditionnal probability that the roi is higher than 0.1 knowing that the loan will be charged off

```{r}
proba_rentable <- data_roi %>% filter(real_roi>0.1) %>% summarise(prob_rentable=n()/nrow(data_roi))
proba_default <- data_roi %>% filter(loan_status=="Charged Off") %>% summarise(prob_default=n()/nrow(data_roi))

proba_default_sachant_rentable <- data_roi %>% filter(loan_status=="Charged Off", real_roi>0.1 ) %>% summarise(prob_default=n()/nrow(data_roi %>% filter(real_roi>0.1)))
  

proba_rentable_sachant_default <- data_roi %>% filter(loan_status=="Charged Off", real_roi>0.05 ) %>% summarise(prob_default=n()/nrow(data_roi %>% filter(loan_status=="Charged Off")))
# no output ??
```





